---
title: "Assignment 3 - 607"
author: "Evelyn Bartley"
date: "2024-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Overview:

Assignment 3 asks me to create a .csv file that includes specific information regarding two airlines (Alaska and AM West) across five destinations (Los Angeles, Phoenix, San Diego, San Francisco, and Seattle). I will read the file into R and perform analysis to compare arrival delays for the two airlines.

I created a .csv file from a google sheet document, saved it in my github repository, copied the url of the raw file, and pasted it into R.

```{r}
#Load the .csv file into R and save it
dataF <- read.csv('https://raw.githubusercontent.com/evelynbartley/Data-607/main/Untitled%20spreadsheet%20-%20Sheet1.csv', header = TRUE)
#fill in empty cells
dataF$X = c("Alaska", "Alaska", "AM West", "AM West")
#replace periods with a space for aesthetic purposes
dataF <- dataF |>
  rename(
    "Los Angeles" = "Los.Angeles",
    "San Diego" = "San.Diego",
    "San Francisco" = "San.Francisco"
  )
dataF
```

## 1. TIDY!

I want to make every column a variable, and every variable a column every observation a row and every row an observation, and every value a cell and every cell a value.

This dataset has 4 variables: Airline, Arrival Status, Destination and Count

Airline has two possible values: Alaska or AM West

Arrival Status has two possible values: on time or delayed

Destination has five possible values: Los Angeles, Phoenix, San Diego, San Francisco, or Seattle

Counts is a quantitative value

```{r}
#Use pivot_longer() to combine the five destination columns into one column that specifies destination, and put those values into a frequency column that specifies the frequency of the observation.
dataF <- dataF |>
  pivot_longer('Los Angeles':'Seattle', names_to = "Destination", values_to = "Counts")
head(dataF)
```

I want to give each column a descriptive name.

```{r}
dataF <- dataF |>
  rename(
    "Airline" = "X",
    "Status" = "X.1"
  )
head(dataF)
```

Now I have a tidy data frame since each variable forms a column, each observation forms a row, and each value has its own cell.

## 2. Analysis

I want to compare the arrival delays for each airline. Let's start by seeing some summary statistics.

```{r}
#Create a subset of the dataframe that only includes information from Alaska airlines
AlaskaData <- dataF[which(dataF$Airline=='Alaska'), ]
AlaskaData
#Create a subset of the dataframe that only includes information from Alaska airlines
AMWestData <- dataF[which(dataF$Airline=='AM West'), ]
AMWestData
#get the total number of planes from each airline
TotalAlaska <- summarise(AlaskaData, sum(Counts))
TotalAlaska
TotalAMWest <- summarise(AMWestData, sum(Counts))
TotalAMWest
```

In total, Alaska airlines had 3,775 flights and AM West had 7,225 flights.

Now let's group the data frame by arrival status.

```{r}
delay <- group_by(dataF, Status)
summarise(delay, 
          sum(Counts))
```

In total, 1,288 flights were delayed and 9,712 were on time.

I want to see the proportion of delayed/on time flights based on what airline they are from.

```{r}
propStatusAirline <- ggplot(dataF, mapping = aes(x = Airline, y = Counts, fill = Status)) + geom_col(position = "fill") 
propStatusAirline
```

Based on this stacked bar chart that shows proportions, there is not a huge difference in the proportion of delayed flights from Alaska vs from AM West.

Now I want to see if there's a big difference in proportion of delayed flights based on their destination for Alaska flights.

```{r}
propAlaskaDestination <- ggplot(AlaskaData, mapping = aes(x = Destination, y = Counts, fill = Status)) + geom_col(position = "fill")
propAlaskaDestination
```

I dont like how the names of the destinations are hard to read so I want to make this proportional bar plot horizontal!

```{r}
propAlaskaDestinationH <- ggplot(AlaskaData, mapping = aes(x = Destination, y = Counts, fill = Status)) + geom_col(position = "fill") + coord_flip()
propAlaskaDestinationH
```

Much better!

Now lets do the same for the AM West flights. I want to see if there's a big difference in proportion of delayed flights based on their destination for AM West flights.

```{r}
propAMWestDestination <- ggplot(AMWestData, mapping = aes(x = Destination, y = Counts, fill = Status)) + geom_col(position = "fill") + coord_flip()
propAMWestDestination
```

## Conclusions

Based on this analysis, there is not a significant difference between arrival delays for Alaska and AM West. From this data specifically, I found that the greatest proportion of delayed flights were from AM West airlines going to San Francisco. I also found that the greatest proportion of on time flights were from Alaska airlines going to Phoenix.
