---
title: "Project 1 - 607"
author: "Evelyn Bartley, AJ Strauman-Scott, James Naval"
date: "2024-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(stringr)
```

## Overview: Project 1 Data 607
In this project, we take a text file

```{r}
chess <- read_delim("https://raw.githubusercontent.com/evelynbartley/Data-607/main/Chess%20File", delim = "|", col_names = c("state", "name", "number", "pointTotal", "prerating", "avgOpp"), skip = 4, show_col_types = FALSE)
```
```{r}
# Convert the file to a data frame
chess <- data.frame(chess)
```

```{r}
# Remove the NA rows
chess <- subset(chess, chess$state != chess$state[3])
```



Here we use an iteration to create a data frame from the odd rows and a separate data frame for the even rows, and then horizontally merge those two data frames to create a data frame with 22 variables and 64 observations.
```{r}
name_rows <- data.frame()
state_rows <- data.frame()
# Use recursion to create two data frames
for (i in 1:nrow(chess)) {
  this_row <- chess[i,]
  if (i %% 2 == 1) {
    name_rows <- rbind(name_rows, this_row)
  } 
  else {
    state_rows <- rbind(state_rows, this_row)
  }
}
#Combine the two data frames
chess1 = data.frame(name_rows, state_rows, check.names=T)
```

Delete the columns with "NA" and other irrelevant columns.
```{r}
chess2 <- select(chess1, -c("X11","X11.1", "number.1", "pointTotal.1", "prerating.1", "avgOpp.1", "X7.1", "X8.1", "X9.1", "X10.1"))
```

Rename the columns.
```{r}
chess3 <- 
rename(chess2,
    "PlayerID" = "state",
    "PlayerName" = "name",
    "TotalPoints" = "number",
    "Round1" = "pointTotal",
    "Round2" = "prerating",
    "Round3" = "avgOpp",
    "Round4" = "X7",
    "Round5" = "X8",
    "Round6" = "X9",
    "Round7" = "X10",
    "State" = "state.1",
    "Prerating" = "name.1"
)
```



Let's make a separate column to isolate the prerating for each player:
```{r}
chess4 <- separate_wider_delim(chess3, cols = "Prerating", delim = ":", names = "NewCol", too_many = "debug")
chess4
chess5 <- separate_wider_delim(chess4, cols = "Prerating_remainder", delim = "->", names = "NewCol2", too_many = "debug")
```


Here we get rid of the irrelevant columns.  
```{r}
chess6 <- select(chess5, -c("NewCol", "Prerating","Prerating_ok", "Prerating_pieces", "Prerating_remainder", "Prerating_remainder_ok", "Prerating_remainder_pieces", "Prerating_remainder_remainder"))
```

We rename the column that contains the preratings.
```{r}
chess7 <- rename(chess6,
                 "PlayerPreRating" = "NewCol2")
```

We get rid of all the ":"'s that precede our preratings. And expand the players preratings that end with a "P" and a number. 
```{r}
chess7$PlayerPreRating <- gsub(":","",as.character(chess7$PlayerPreRating))

chess8 <- separate_wider_delim(chess7, cols = "PlayerPreRating", delim = "P", names = "PreRating", too_many = "debug")
```
Clean it up.
```{r}
chessF <- select(chess8, -c("PlayerPreRating", "PlayerPreRating_ok", "PlayerPreRating_pieces", "PlayerPreRating_remainder"))
```


Here we isolate the player ID's for each round by removing the letter that represents the result of the round.
```{r}
chessF$Round1 <- gsub('[A-Z]', "", chessF$Round1)
chessF$Round2 <- gsub('[A-Z]', "", chessF$Round2)
chessF$Round3 <- gsub('[A-Z]', "", chessF$Round3)
chessF$Round4 <- gsub('[A-Z]', "", chessF$Round4)
chessF$Round5 <- gsub('[A-Z]', "", chessF$Round5)
chessF$Round6 <- gsub('[A-Z]', "", chessF$Round6)
chessF$Round7 <- gsub('[A-Z]', "", chessF$Round7)
```

Now we want to replace each player id that shows in each round by the corresponding player id's prerating.

I want to change every row with numbers to a numeric vector.
```{r}
chessF$PlayerID = as.numeric(sub("[A-Z]", "", chessF$PlayerID))
chessF$Round1 = as.numeric(sub("[A-Z]", "", chessF$Round1))
chessF$Round2 = as.numeric(sub("[A-Z]", "", chessF$Round2))
chessF$Round3 = as.numeric(sub("[A-Z]", "", chessF$Round3))
chessF$Round4 = as.numeric(sub("[A-Z]", "", chessF$Round4))
chessF$Round5 = as.numeric(sub("[A-Z]", "", chessF$Round5))
chessF$Round6 = as.numeric(sub("[A-Z]", "", chessF$Round6))
chessF$Round7 = as.numeric(sub("[A-Z]", "", chessF$Round7))
chessF$PreRating = as.numeric(sub("[A-Z]", "", chessF$PreRating))
```


This for loop replaces the opponent's ID with their prerating in Round 1.
```{r}
for (i in chessF$Round1)
  {
  chessF$Round1[i] <- chessF$PreRating[chessF$Round1[i]]
}
for (i in chessF$Round2)
  {
  chessF$Round2[i] <- chessF$PreRating[chessF$Round2[i]]
}
for (i in chessF$Round3)
  {
  chessF$Round3[i] <- chessF$PreRating[chessF$Round3[i]]
}
for (i in chessF$Round4)
  {
  chessF$Round4[i] <- chessF$PreRating[chessF$Round4[i]]
}
for (i in chessF$Round5)
  {
  chessF$Round5[i] <- chessF$PreRating[chessF$Round5[i]]
}
for (i in chessF$Round6)
  {
  chessF$Round6[i] <- chessF$PreRating[chessF$Round6[i]]
}
for (i in chessF$Round7)
  {
  chessF$Round7[i] <- chessF$PreRating[chessF$Round7[i]]
}
head(chessF)
```



This line of code takes the seven opponents prerating numbers and take the average of each row and store them in a new column. 
```{r}
chessF <- chessF |>
  rowwise() |>
  mutate('OppAvgPrerating' = mean(c(Round1, Round2, Round3, Round4, Round5, Round6, Round7), na.rm = TRUE))

chessF <- chessF |>
  rowwise() |>
  mutate('OppAvgPrerating' = round(OppAvgPrerating, digits = 0))

```

##Conclusion: I am left with a dataframe with information on Player's Name, Total Points, State, PreRating, and Average Pre Chess Rating of Opponents for each player.

```{r}
chessF <- select(chessF, -c("PlayerID", "Round1", "Round2", "Round3", "Round4", "Round5", "Round6", "Round7"))

head(chessF)
```
Here is the .csv file:
```{r}
write_csv(chessF, 'chessFinal.csv')
```

