---
title: "Project 2 - part 2"
author: "Evelyn Bartley"
date: "2024-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Overview

I am taking the .csv file of the Annual Surface Tempurature Change from climatedata.imf.org. My goal is to tidy the dataset for analysis. I want to see which regions are most affected by climate change.

Lets read the .csv file into R from my github repository.
```{r}
climate <- read.csv("https://raw.githubusercontent.com/evelynbartley/Data-607/main/Indicator_3_1_Climate_Indicators_Annual_Mean_Global_Surface_Temperature_577579683071085080.csv")
tibble(climate)
```


Let's clean up the dataset to include the variables we need for analysis. I want to use the ISO3 code for each country instead of the Country's name for tidyness.
```{r}
climate1 <- climate |> 
  select(Country = ISO3, X1961:X2000)
tibble(climate1)
```


Instead of having columns for every year, I want to have one column for year, and one column for the surface temperature change in degrees Celcius.
```{r}
climate2 <- climate1 %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "Year",
    values_to = "TempChange"
  )
tibble(climate2)
```
To create one value that we can reference for the change in surface temperature from 1961 to 2000, I want to calculate the average change in surface temperature for each country.
```{r}
climate3 <- climate2 %>%
  group_by(Country) %>%
  summarise(avg = mean(TempChange, na.rm = TRUE))
head(climate3)
```

Lets see the distribution of average change in temperature.
```{r}
avgofavgs <- mean(climate3$avg, na.rm = TRUE)
ggplot(climate3, aes(x = avg)) + geom_histogram() + geom_vline(aes(xintercept = avgofavgs), color = "tomato", linewidth = 1)
```
Our distribution is looking pretty normal! There does seem to be two outliers.

I want to see which country had the highest average change in temperature and which had the lowest.
```{r}
climate3[which.min(climate3$avg), ]
climate3[which.max(climate3$avg), ]
```
## Analysis and Conclusion

The country with the lowest average change in temperature was Greenland at -0.156 degrees Celcius and the country with the highest average change in temperature was Luxembourg at 1.651 degrees Celcius.
Using the small arrows when viewing climate3, we can arrange the averages in descending order. From this, I can see that Luxenbourg, Belgium, Estonia, Latvia, and Slovenia had the five highest average changes in temperatures. This provides evidence that Europe is being effected by climate change the most. 

