---
title: "Web APIs Assignment - 607"
author: "Evelyn Bartley"
date: "2024-03-18"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

## Overview:

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(httr2)
library(jsonlite)
```


After creating an account and signing up for an API key, I chose the Most Popular API from the New York Times APIs.
My goal is to use the httr2 package to read in the JSON data.

```{r}
#request the most viewed articles for the last 7 days
#https://api.nytimes.com/svc/mostpopular/v2/viewed/{period}.json
request <- request('https://api.nytimes.com') %>%
  req_url_path_append(
    'svc',
    'mostpopular',
    'v2',
    'viewed',
    '7.json?api-key=gAEU2kmMwr0EEG1gwEzF9YzKf0lfClsc'
  )

nyt_resp <- req_perform(request)

```

```{r}
nyt_resp %>% resp_content_type()
#the response is an application/json
```

```{r}
#nyt_resp %>% resp_body_json()
#confirmed the response body is in json format
# Parse the JSON response
nyt_data <- nyt_resp %>%
  resp_body_json(check_type = TRUE, simplifyVector = FALSE)
```

```{r}
# Extract from article
articles <- nyt_data$results$title
nyt_data$results$source
```

This first approach wasn't successful. I could look at the api response data of the application/json format but couldn't extract elements from it.


Using the tip about the jsonlite package Amanda Fox gave me on slack:

```{r}
my_url <- 'https://api.nytimes.com/svc/mostpopular/v2/viewed/7.json?api-key=gAEU2kmMwr0EEG1gwEzF9YzKf0lfClsc'
articledata <- fromJSON(my_url, flatten = TRUE)
title <- articledata$results$title
author <- articledata$results$byline
date_published <-  articledata$results$published_date 
section <- articledata$results$section
subsection <- articledata$results$subsection

articledf <- data.frame(title, author, date_published, section, subsection)
tibble(articledf)
```

I looked into using the nytimes package to do this assignment but I wasn't able to get anywhere.

```{r}
library(nytimes)
nytimes_key("gAEU2kmMwr0EEG1gwEzF9YzKf0lfClsc")
```

```{r}
viewed <- ny_popular_viewed(7)
```


## Conclusion:
I was surprised that I could have completed the assignment with just the one chunk of code. Messing around with the req_*() and resp_*() functions did help me understand apis more though!
